Freezing of a Weeks-Chandler-Anderson Fluid using classical Density Functional Theory

### 2020-11-08 papers/thesis-kirstie/thesis.tex
edits

A description of cDFT ordinarily begins with considering the Grand Free 
Energy defined as
\begin{equation}\Phi=F-\mu{N}\end{equation}
%\begin{equation}\int\frac{N}{V}\frac{V_{ext}}{N}dV\end{equation}
which can then be expressed in terms of a varrying number density 
profile $n(\vec r)$
%\begin{equation}\Phi=F-\int n(\vec{r})\phi{(\vec r)}d\vec{r} +\int n(\vec{r})\phi{(\vec r)}d\vec{r}-\mu{N}\end{equation}
\begin{equation}\label{GrandFE}\Phi= F_{int} +\int n(\vec{r})\phi{(\vec r)}d\vec{r}-\mu\int n(\vec r)d\vec{r}\end{equation}
where $\phi=\frac{V_{ext}}{N}$, $V_{ext}(\vec r)$ is an externally 
applied potential energy, and the Helmholtz free energy is expressed as
\begin{equation}F = F_{int} + \int n(\vec{r})\phi{(\vec r)}d\vec{r}\end{equation}  
%If $\mu=0$ and $V_{ext}=0$, then the Grand free energy $\Phi$ reduces 
%to the Helmholz free energy $F$. %The second and third terms in Eq~\ref{GrandFE} 
%are both functionals of $n(\vec{r})$, so it remains to show that $F$, and 
%thus $\Phi$, is a functional of $n(\vec{r})$.

@@

The functional for the ideal Helmholtz free energy is given by 
Equation~\ref{eq:Fideal-n(r)}, and the functional for the excess Helmholtz 
free energy depends on the particular fluid as defined by the spheres' 
interatomic potential V(r). To implement cDFT, the number density profile 
$n(\vec{r})$ is varied with the temperature fixed until the free energy 
is minimized. 
\begin{equation}{F[n(\vec{r})]=F_{ideal}[n(\vec{r})] + F_{excess}[n(\vec{r})]}\end{equation}
The minimum free energy and its corresponding number density 
profile are then the equilibrium values for that temperature.

\section{Soft Fundamental Measure Theory (SFMT)}
In order to use cDFT, a functional for the excess Helmholtz free
energy for the particular fluid being modeled must be formed. But forming 
an expression based on Equation~\ref{eq:Fexcess-simplified} can be challenging. 
For the Hard Sphere fluid, there is another way to form an expression for 
the excess Helmholtz free energy in terms of a homogeneous number density 
and, that is to use Scaled Particle Theory (SPT)~\cite{ReissSPT}.

>>>>

A description of cDFT often begins with considering the Grand Free 
Energy defined as
\begin{equation}\Phi=F-\mu{N}\end{equation}
%\begin{equation}\int\frac{N}{V}\frac{V_{ext}}{N}dV\end{equation}
which can then be expressed in terms of a varrying number density 
profile $n(\vec r)$
%\begin{equation}\Phi=F-\int n(\vec{r})\phi{(\vec r)}d\vec{r} +\int n(\vec{r})\phi{(\vec r)}d\vec{r}-\mu{N}\end{equation}
\begin{equation}\label{GrandFE}\Phi= F_{int} +\int n(\vec{r})\phi{(\vec r)}d\vec{r}-\mu\int n(\vec r)d\vec{r}\end{equation}
where $\phi=\frac{V_{ext}}{N}$ 
is the externally 
applied potential energy $V_{ext}(\vec r)$ per particle, 
and the Helmholtz free energy is expressed as
\begin{equation}F = F_{int} + \int n(\vec{r})\phi{(\vec r)}d\vec{r}\end{equation}  
%If $\mu=0$ and $V_{ext}=0$, then the Grand free energy $\Phi$ reduces 
%to the Helmholz free energy $F$. %The second and third terms in Eq~\ref{GrandFE} 
%are both functionals of $n(\vec{r})$, so it remains to show that $F$, and 
%thus $\Phi$, is a functional of $n(\vec{r})$.

@@

%According to classical Density Functional Theory,

%A number density profile is a time-invariant average 
%of the spacial distribution of the particles that make up the fluid, and
%There will be one, and only one 
%that minimizes the Helmholtz free energy.
%The minimum free energy, 
%is the value of the Helmholtz free energy that
%the system will have when it is in equilibrium at that temperature, 
%and its corresponding number density profile 
%are the equilibrium values the system will have at that temperature.
%is the equilibrium number density profile. 
%
In this section 
%To form the functional for the Helmholtz free energy, expressions for 
a functional for the Helmholtz free energy was formed by adding 
the ideal and the excess portions of the free 
energy.
\begin{equation}{F[n(\vec{r})]=F_{ideal}[n(\vec{r})] + F_{excess}[n(\vec{r})]}\end{equation}
The functional for the ideal portion of the Helmholtz free energy is given by 
Equation~\ref{eq:Fideal-n(r)}, and a functional for the excess portion of the Helmholtz 
free energy is given in Equation~\ref{eq:Fexcess-simplified}. The specific 
form for the excess Helmholtz free energy depends on the particular 
fluid being modeled as defined by the  
interatomic potential energy V(r) between the spheres, which can make
forming an expression difficult. 
Once an expression for the Helmholtz free energy as a functional of the
number density profile $n(\vec r)$ is formed then, according to classical 
Density Functional Theory, the equilibrium value of the 
Helmholtz energy, and its cooresonding equilibrium density profile can be
found by varying the number density profile
%can be varied 
through all possible configurations at some fixed temperature
until the free energy is minimized.

\section{Soft Fundamental Measure Theory (SFMT)}
In order to use classical Density Functional Theory, a functional for 
the excess Helmholtz free
energy for the particular fluid being modeled must be formed. But forming 
an expression based on Equation~\ref{eq:Fexcess-simplified} can be challenging. 
For the Hard Sphere fluid, there is another way to form an expression for 
the excess Helmholtz free energy in terms of a homogeneous number density 
and, that is to use Scaled Particle Theory (SPT)~\cite{ReissSPT}.
